---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Data pre-processing
Loading the data
```{r cache=TRUE}
activity <- read.csv("./activity.csv")
```
Explore the data
```{r}
summary(activity)
str(activity)
head(activity)
```
Count number of records with missing data
```{r}
sum(is.na(activity$steps))
```

Calculate the total number of steps taken per day
```{r}
DailyTotalSteps <- tapply(activity$steps, activity$date, sum)
```
Visualize the trend
```{r}
p <- barplot(DailyTotalSteps, xlab="Date", ylab="total number of steps taken per day")
```

Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(DailyTotalSteps, na.rm=TRUE); median(DailyTotalSteps, na.rm=TRUE)
```
The mean total number of steps per day is 10,766, whereas the median is 10,765.

Compute the average number of steps for every interval across Oct. and Nov.
```{r}
meanStepsByInterval <- tapply(activity$steps, activity$interval, na.rm = TRUE, mean)
```
Visualize the trend
```{r}
g <- plot(meanStepsByInterval ~ names(meanStepsByInterval), type = "l", xlab="5-min Interval/day", ylab="mean number of steps taken per 5-min interval")
```

Check when average number of steps peaks.
```{r}
peakElement <- which.max(meanStepsByInterval)
names(meanStepsByInterval)[peakElement]
```
The step activities peak at around 8:35 am in the morning.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
